---
title: "Iowa Caucus results 2020"
author: "Desi Quintans"
date: "Compiled `r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

librarian::shelf(tidyverse, rvest)
```

```{r}
# Some utility functions

drop_blank <- function(vec) {
    vec[str_length(vec) > 0]
}

strip_html <- function(vec, replace = ",") {
    str_replace_all(vec, "(<.*?>\\n<.*?>)|(<.*?>)", ",")
}

omit_element <- function(vec, regex) {
    vec[str_detect(vec, regex) == FALSE]
}

`%notin%` <- function(x, y) {
    !(x %in% y)
}
```


# Aim

To turn the Iowa Caucus' HTML page into a dataframe.



# Generate a snapshot of the current results

## Import data

```{r}
compile_time <- format(Sys.time(), '%Y-%m-%dT%H%MZ+10')
```


The time of access is: `r compile_time`.

```{r}
iowa <- read_html("https://results.thecaucuses.org/")
```

The HTML file is coded as a list (`<ul>`) that is styled using other technology, it is not coded as a table. Makes it harder to access the information.



## Build the table header

```{r}
# The list of candidates comes from the true table header.
candidates <- 
    iowa %>% 
    html_nodes("ul.thead li") %>% 
    html_text() %>% 
    drop_blank() %>% 
    omit_element("County|Precinct")

candidates


# The list of measures comes from the table's subheader.
table_subhead <- 
    iowa %>% 
    html_nodes("ul.sub-head li") %>% 
    html_text() %>% 
    drop_blank()
    
table_subhead


# The two can be combined so that the dataframe has only one header row.
table_header <- 
    paste(rep(candidates, each = 3), table_subhead, sep = "_") %>% 
    str_remove_all(" Expression")

table_header
```


## Which precincts belong to which counties?

```{r}
county_names <- 
    iowa %>% 
    html_nodes("div.precinct-rows div.precinct-county div.wrap") %>% 
    html_text()

county_names


# This is a list of 104 nodes, so each element corresponds to each county.
precinct_data <- 
    iowa %>% 
    html_nodes("div.precinct-rows div.precinct-data")


# This is the number of rows per county. Note that it is not the number of 
# *precincts* per county, because it includes the 'total' row for now.
county_rowcount <- 
    precinct_data %>% 
    as.character() %>% 
    str_count("<li>.*?[A-Z].*?</li>")

county_rowcount


# The name of each row, recognised as a list item that has capital letters in 
# it and not just numbers.
precinct_names <- 
    precinct_data %>% 
    as.character() %>% 
    str_extract_all("<li>.*?[A-Z].*?</li>", simplify = FALSE) %>% 
    flatten_chr() %>% 
    str_remove_all("<.*?>")


# A lookup table for matching precincts to counties.
precinct_table <- 
    tibble(county = rep(county_names, times = county_rowcount),
           precinct = precinct_names)

head(precinct_table, 20)
```


## Grab SDE counts per precinct

```{r}
precinct_counts <- 
    precinct_data

head(precinct_counts)

precinct_data[[1]]
```


## Format SDE counts as a table

```{r}
precinct_counts <- 
    precinct_data %>% 
    as.character() %>% 
    # Every precinct in a county is currently on a single line. Split each 
    # precinct into its own element.
    str_remove_all("\\n") %>% 
    str_split("<ul.*?>") %>% 
    # Flatten into an atomic Char vector and remove non-data HTML elements
    flatten_chr() %>% 
    str_remove_all("<li>|</div>|</ul>") %>% 
    omit_element("^<div") %>% 
    # Format each line for reading-in as delimited values. Values are delimited
    # with ? because both , and ; are used in precinct names.
    str_replace_all("</li>", " ? ") %>% 
    str_remove("\\? $") %>% 
    # Create a dataframe from it
    read_delim(., delim = "?", col_names = prepend(table_header, "precinct")) %>% 
    mutate_all(str_trim)

head(precinct_counts)
```


## Merge the counts with the counties

```{r}
# If the precinct names are the same and occur in the same order, then the 
# two dataframes can simply be column-bound.
if (identical(precinct_table$precinct, precinct_counts$precinct)) {
    clean_results <- 
        bind_cols(precinct_table, select(precinct_counts, -precinct)) %>%
        mutate_at(vars(-(county:precinct)), parse_number) %>% 
        mutate_at(vars(matches("First|Final")), as.integer)
} else {
    stop("The row order of 'precinct_table' and 'precinct_counts' is not identical.")
}

glimpse(clean_results)
```



## Pivot into long format


```{r}
tidy_result <- 
    clean_results %>% 
    pivot_longer(-(county:precinct)) %>% 
    separate(name, into = c("candidate", "measure")) %>% 
    mutate(measure = recode(measure, 
                            First = "expression_1", 
                            Final = "expression_2",
                            SDE = "sde"))
```


## Recalculate totals

```{r}
new_totals <- 
    tidy_result %>% 
    filter(precinct != "Total") %>% 
    group_by(county, candidate, measure) %>% 
    summarise(Total_recalc = sum(value)) %>% 
    ungroup() %>% 
    pivot_longer(Total_recalc, names_to = "precinct", values_to = "value")


recalculated <- 
    full_join(tidy_result, new_totals) %>% 
    arrange(county, precinct, candidate, measure)
```


## Are there any totalling errors?

```{r}
recalculated %>% 
    filter(str_detect(precinct, "Total")) %>% 
    pivot_wider(names_from = "precinct", values_from = "value") %>% 
    filter(near(Total, Total_recalc) == FALSE) %>% 
    print()
```


## Output

```{r}
write_csv(recalculated, paste0("_compiled/", compile_time, ".csv"))

recalculated
```



# Pairwise comparison of all results

It is possible that snapshots that were downloaded hours apart have no differences between them. Which snapshots should be looked at, and which ones are unchanged and can be ignored?

## Import all snapshots

```{r}
snapshots <- list.files("_compiled", full.names = TRUE)

snapshots

snapshot_dfs <- map(set_names(snapshots), 
                    ~ read_csv(.x, col_types = "ccccd"))
```



## Work out combinations

```{r}
perms <- 
    combn(snapshots, m = 2, simplify = FALSE)

perm_names <- map_chr(perms, ~ paste(..1[1], ..1[2], sep = " == "))

perm_names
```


## Are the dataframes identical?

```{r}
pairwise_identical <- 
    map(perms, ~ identical(pluck(all_snapshots, ..1[[1]]),
                           pluck(all_snapshots, ..1[[2]]))) %>% 
    set_names(perm_names)

pairwise_identical
```

